---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

### 1. Load the data

```{r}
activity <- read.csv(unz("activity.zip", "activity.csv"))
head(activity)
tail(activity)
summary(activity)
```

### 2. Process/transform the data (if necessary) into a format suitable for your analysis

#### No transformation necessary at this point.

## What is mean total number of steps taken per day?

### 1. Make a histogram of the total number of steps taken each day
```{r}
steps_date <- activity[ , colnames(activity) %in% c('steps','date') ]
steps_dsum <- aggregate(.~date, data=steps_date, sum)
hist(steps_dsum[,2], 
     xlab="Steps",
     ylim=(c(0,40)),
     col='light blue',
     main="Histogram of Total Number of Steps Each Day")
```

### 2. Calculate and report the **mean** and **median** total number of steps taken per day
```{r}
mean(steps_dsum$steps, na.rm=TRUE)
median(steps_dsum$steps,na.rm=TRUE)
```

## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps_interval <- activity[ , colnames(activity) %in% c('steps','interval') ]
steps_imean <- aggregate(.~interval, data=steps_interval, mean)
head(steps_imean)
summary(steps_imean)
library("lattice")
xyplot(steps ~ interval,
       data = steps_imean,
       type = "l",
       lty = 1,
       lwd = 1,
       main = "Time Series of Average Number of Steps Across All Days",
       xlab = "5-Minute Interval",
       ylab = "Average Number of Steps Across All Days",
       col.line = "red")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
steps_imean[steps_imean$steps == max(steps_imean$steps), ]
```

## Imputing missing values

### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r}
a <- table(complete.cases(activity))
a[names(a)==FALSE]
```

### 2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

#### We decide to fill each missing value with the mean for that 5-minute interval

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_complete <- activity
na.steps <- which(is.na(activity_complete$steps))
na.interval <- activity_complete$interval[na.steps]
activity_complete$steps[na.steps] <- steps_imean$steps[match(na.interval, steps_imean$interval)]
b <- table(complete.cases(activity_complete))
b[names(b)==FALSE]
head(steps_imean)
head(activity_complete)
tail(steps_imean)
tail(activity_complete)
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
steps_date_complete <- activity_complete[ , colnames(activity_complete) %in% c('steps','date') ]
steps_dsum_complete <- aggregate(.~date, data=steps_date_complete, sum)
hist(steps_dsum_complete[,2], 
     xlab="Steps", 
     ylim=(c(0,40)),
     col='light blue',
     main="Histogram of the Total Numer of Steps")
mean(steps_dsum_complete$steps, na.rm=TRUE)
median(steps_dsum_complete$steps,na.rm=TRUE)
```

#### The above hisgram graph shows that the maximum frequecy is increased from (this makes sense since the the NA values are now replaced by the mean step values for the same 5-minute interval. The number of steps for those intervals must be increased (. However, the mean total number of steps taken per day has not changed.  This is because we replaced the NA with the mean number of steps for that interval. The only estimate changed is the median of total number of steps taken per day -- it is now the same as the mean (10766.19).


## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
day <- weekdays(as.Date(activity_complete$date))
wkday_list <- which(day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
wkend_list <- which(day %in% c("Saturday","Sunday"))
day <- replace(day, wkday_list, "weekday")
day <- replace(day, wkend_list, "weekend")
activity_complete$weekday <- day
head(activity_complete)
tail(activity_complete)
```

### 2. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
df_weekday <- activity_complete[activity_complete$weekday == 'weekday', ]
steps_interval_weekday <- df_weekday[ , colnames(df_weekday) %in% c('steps','interval') ]
steps_weekday_mean <- aggregate(.~interval, data=steps_interval_weekday, mean)
steps_weekday_mean$weekday <- rep('weekday',length(steps_weekday_mean$steps))

df_weekend <- activity_complete[activity_complete$weekday == 'weekend', ]
steps_interval_weekend <- df_weekend[ , colnames(df_weekend) %in% c('steps','interval') ]
steps_weekend_mean <- aggregate(.~interval, data=steps_interval_weekend, mean)
steps_weekend_mean$weekday <- rep('weekend',length(steps_weekend_mean$steps))

summary(steps_weekday_mean)
summary(steps_weekend_mean)

steps_mean <- rbind(steps_weekday_mean, steps_weekend_mean)
xyplot(steps ~ interval | weekday,
       data = steps_mean,
       type = "l",
       lty = 1,
       lwd = 1,
       #main = "Time Series of Average Number of Steps",
       xlab = "5-Minute Interval",
       ylab = "Average Number of Steps",
       col.line = "red",
       layout=c(1,3)
       )
```

#### The above graphics and statistics show that there are differences in activity patterns between weekdays and weekends. The mean and median of total number of steps in weekend are higher than the weekdays. The maximum number of steps of weekend is lower than that of the weekends. The activities over weekend are more steady than the weekdays.

